%
\chapter*{Acknowledgments}
%

%
\subsectionend $ $\\
%
\emph{The mathematician does not study pure mathematics because it is useful;
he studies it because he delights in it and he delights in it because it
is beautiful.} \\
%
\ib{-- ~ Henri Poincar\'e}
%
\subsectionend $ $\\
%

I would first like to thank my family for their unceasing love and support.
You all have provided a warm and nurturing environment for me to develop into
who I am today. Without you, none of this would have been possible.

I would also like to thank my DRP mentor, Arun Debray. It has been an absolute
pleasure reading with you, and I have learned an enormous amount of mathematics
over the course of all our conversations. Your help in writing this thesis
has been invaluable, and your puns and jokes will be missed.

To all my friends -- you guys are the reason that my undergraduate experience
has been so wonderful. I've loved staying up late doing homework with you and
all the conversations we've had about mathematics. I hope our mathematical
careers will continue to intersect in the upcoming years, and I wish you
the best in your graduate studies, and in any other pursuits.

My interest in pure mathematics was almost non-existent before I started my
undergraduate career here at the University of Texas at Austin. I'd like to
thank the mathematics department here for providing an environment for me
appreciate mathematics and grow into an aspiring mathematician. In particular,
I'd like to thank Professor Andy Neitzke for his invaluable advice and mentorship,
you are one of the reasons I fell in love with geometry here in Austin.
I would also like to thank Professors Sean Keel and Cameron Gordon for teaching some
of my favorite classes, and Professor Daniel Allcock for the many wonderful conversations
I've had from popping in unannounced into his office.

Last but not least, I would like to thank Professor Dan Freed for supervising
this senior thesis, in addition to all of his guidance during my time here at UT.
You have been and inexhaustible source of advice and insight, and it is safe to
say that I would not be anywhere near where I am now without your mentorship.
You have been an enormous influence on the kind of mathematics I want to
pursue, and have played an integral part in shaping how I view geometry
and topology. You are one of my mathematical role models, and I would be extremely
lucky to even be an infinitesimal affine approximation of you.
%