%
\section{Dirac Operators in $\R^n$}
%
One important application of Clifford algebras and Spin groups comes from
physics. In the process of developing a relativistic equation for the electron,
Paul Dirac saw the need for a first order differential operator $D$ such
that $D$ squared to the Laplace operator\footnote{Our reasoning for choosing this sign
convention for $\Delta$ is twofold -- one reason is that the spectrum of $\Delta$
is positive with this choice of sign, and the second is that this definition
coincides with a generalized Laplace operator on a Riemannian manifold.}.
\[
\Delta = -\sum_i \frac{\partial^2}{\partial(x^i)^2}
\]
If $D$ were the be first order, it would have to be written as
\[
D = a^i \frac{\partial}{\partial x^i}
\]
for some coefficients $a^i$. However, it is clear that choosing scalar coefficients
for the $a^i$ will not suffice. For example, in $\R^2$ with the standard coordinates
$x$ and $y$, any first order operator $D = a^1\partial_x + a^2\partial_y$ satisfies
\[
D^2 = \left(a^1\frac{\partial}{\partial x} + a^2\frac{\partial}{\partial y}\right)^2
= (a^1)^2\frac{\partial^2}{\partial x^2} + a^1a^2\frac{\partial^2}{\partial x \partial y}
+ a^2a^1 \frac{\partial^2}{\partial y \partial x} + (a^2)^2 \frac{\partial^2}{\partial y^2}
\]
From this equation, we see that the $a^i$ must square to $-1$, and we must also
have that $a^1a^2 + a^2a^1 = 0$ in order for the mixed partial terms to vanish.
This is not possible if the $a^i$ are scalars (in either $\R$ or $\C$).
However, the required relations are exactly the relations between orthogonal
basis vectors in the Clifford algebra $\Cliff_{0,n}(\R)$!
%
\begin{defn}
Let $\set{e_i}$ be the standard basis for $\R^n$, $\set{e^i}$
\footnote{The choice of using the dual basis elements $e^i$ instead of the
$e_i$ is so that $D$ behaves tensorially with respect to coordinate change.}
its dual basis, and $x^1, \ldots ,x^n$ the standard coordinates on $\R^n$.
The \ib{Dirac operator} on $\R^n$ is the first order differential operator
\[
D = e^i\frac{\partial}{\partial x^i}
\]
\end{defn}
%
It is not clear from the definition what function space $D$ should act on.
The partial derivative operators make sense for any vector valued function, but
multiplcation by $e^i$ does not make sense an arbitrary vector space -- it must
be a Clifford module. Therefore, $D$ is an operator
$D: C^\infty(\R^n, M) \to C^\infty(\R^n, M)$ on smooth functions from $\R^n$
to a Clifford module $M$. In fuller generality, functions from $\R^n \to M$
are equivalent to smooth sections of $\R^n \times M \to \R^n$, which is a bundle
of Clifford modules over $\R^n$. This viewpoint generalizes more naturally,
and the Dirac operator defined on a general Riemannian manifold is an operator
on sections of some bundle. \\

The Dirac operators in dimensions $1$, $2$, and $4$ exhibit some extremely
interesting behavior, corresponding to the appearances of $\C$ and $\H$
when $\R^n$ is given a negative definite bilinear form. In $1$ dimension,
the Clifford algebra $\Cliff_{0,1}(\R)$ admits the ordered basis $(1,e^1)$, which
then gives a basis for $C^\infty(\R, \Cliff_{0,1}(\R))$ as a
$\C^\infty(\R)$ module where for any $f : \R \to \Cliff_{0,1}(\R)$, we have the
decomposition $f(x) = u(x) + e^1 v(x)$ , giving a column vector representation
\[
f = \begin{pmatrix}
u \\
v
\end{pmatrix}
\]
The Dirac operator is $e^1\partial_x$, which is represented in matrix form as
\[
D = \begin{pmatrix}
0 & -\partial_x \\
\partial_x & 0
\end{pmatrix}
\]
Note that the matrix is block off-diagonal. Since we picked the ordered basis $(1,e^1)$.
for $\Cliff_{0,1}(\R)$, whose elements are even and odd respectively. The action
of the top right block is the action on the even subspace, and the action of
the bottom left block is the action on the odd subspace. This
shows that $D$ is an odd operator, as it reverses the grading. \\

In $2$ dimensions, $\Cliff_{0,2}(\R)$ has the ordered basis $(1, e^1e^2, e_1,e_2)$,
where we choose the ordering by the parity of the elements. Again, we can
write any function $f : \R^2 \to \Cliff_{0,2}(\R)$ in terms of this basis
as \[
f(x,y) = f_0(x,y) + e^1e^2f_{12}(x,y) + e^1f_1(x,y) + e^2f_2(x,y)
\]
where the component functions are all elements of $C^\infty(\R^2)$. In this
basis, the action of the Dirac operator $D = e^1\partial_x + e^2\partial_y$
is represented by the matrix of differential operators
\[
D = \begin{pmatrix}
0 & 0 & -\partial_x & -\partial_y \\
0 & 0 & -\partial_y & \partial_x \\
\partial_x & \partial_y & 0 & 0 \\
\partial_y & -\partial_x & 0 & 0
\end{pmatrix}
\]
Which is actually a familiar differential operator in disguise. We make the
identification of $\Cliff_{0,2}(\R)$ with the quaternions $\H$, which
have a direct sum decomposition $\H = \C \oplus \C$, giving a $\Z/2\Z$-grading.
The mappings $e_1e_2 \mapsto i$ and $e_2 \mapsto i$ determine vector space
isomorphisms of the even and odd subspaces of $\Cliff_{0,2}(\R)$ with $\C$,
giving an isomorphism of $\Cliff_{0,2}(\R)$ with $\H = \C \oplus \C$. The
operator in the top right block is then exactly twice the differential operator
\[
-\frac{\partial}{\partial z} = -\frac{1}{2}\left( \frac{\partial}{\partial x}
- i \frac{\partial}{\partial y}\right)
\]
on functions $\R^2 \to \C$, and the operator in the bottom block is the twice
the operator
\[
\frac{\partial}{\partial \overline{z}} = \frac{1}{2}\left( \frac{\partial}{\partial x}
+ i \frac{\partial}{\partial y}\right)
\]
So $D$ is more compactly represented under this identification as
\[
D = \begin{pmatrix}
0 & 2\partial_z \\
2\partial_{\overline{z}} & 0
\end{pmatrix}
\]
In $4$ dimensions, the irreducible module over $\Cliff_{0,3} \cong M_2\H$ is
$\H^2$. This has a direct sum decomposition as $\H^2 = \H \oplus \H$, allowing
us to represent functions $f : \R^4 \to \H^2$ as pairs of functions $\R^4 \to \H$,
which are the maps into the even and odd subspaces of $\H^2 = \H \oplus \H$.
The quaternions admit operators similar to the operators $\partial_z$ and
$\partial_{\overline{z}}$, which are
%
\begin{align*}
\frac{\partial}{\partial q} &= \frac{1}{4}\left( \frac{\partial}{\partial x^1}
- i \frac{\partial}{\partial x^2} - j\frac{\partial}{\partial x^3}
- k \frac{\partial}{\partial x^4}\right) \\[5pt]
\frac{\partial}{\partial \overline{q}} &= \frac{1}{4}\left( \frac{\partial}{\partial x^1}
+ i \frac{\partial}{\partial x^2} + j\frac{\partial}{\partial x^3}
+ k \frac{\partial}{\partial x^4}\right)
\end{align*}
%
Using these identifications, the Dirac operator is compactly represented as
\[
D = \begin{pmatrix}
0 & -4\partial_q \\
4\partial_{\overline{q}} & 0
\end{pmatrix}
\]
which encodes quaternionic analogues of the Cauchy-Riemann equations.
%
%TODO More stuff later, maybe in another section?
%