\chapter{Spin Structures on Manifolds}
%
\subsectionend $ $\\
%
\emph{Mathematics is a part of physics. Physics is an experimental science, a part of
natural science. Mathematics is the part of physics where experiments are cheap.} \\
%
\ib{-- ~ V.I. Arnold}
%
\subsectionend
%TODO motivate some of the definitions (add some flavor)
%
\section{Fiber Bundles}
%
In geometry and topology, we often want to consider families of objects
(e.g. manifolds, vector spaces) that vary smoothly along a manifold like
the tangent spaces of a manifold. This is best formalized in the construction
of a fiber bundle.
%
\begin{defn}
Let $M$ and $F$ be smooth manifolds. Then a \ib{fiber bundle} over $M$ with
model fiber $F$ is a the data of a smooth manifold $E$ with a smooth map
$\pi : E \to M$ such that for every point $p \in M$, there is a neighborhood
$U \subset M$ containing $p$ and a diffeomorphism
$\varphi : \pi\inv(U) \to U \times F$ such that the diagram
\[\begin{tikzcd}
\pi\inv(U) \ar[rr, "\varphi"] \ar[dr, "\pi"']&& U \times F \ar[dl, "p_U"] \\
& U
\end{tikzcd}\]
commutes, where $p_U$ denotes projection onto the first factor. The map $\varphi$ is
called a \ib{local trivialization} of the fiber bundle $\pi : E \to M$. The
manifold $M$ is called the \ib{base space}, while the manifold $E$ is
called the \ib{total space}.
\end{defn}
%
Given a fiber bundle $\pi : E \to M$ we often denote the fiber $\pi\inv(p)$
by $E_p$.
%
\begin{defn}
let $\pi : E \to M$ and $p : L \to M$ be fiber bundles with model fiber $F$
over $M$. A \ib{bundle homomorphism} is a smooth map $\varphi : E \to L$ such
that the following diagram commutes
\[\begin{tikzcd}
E \ar[dr, "\pi"']\ar[rr,"\varphi"] & & L \ar[dl, "p"] \\
& M
\end{tikzcd}\]
\end{defn}
%
An important property of fiber bundles is that they pull back.
\begin{defn}
Let $\pi : E \to M$ be a fiber bundle with model fiber $F$
and let $f : X \to M$ be a smooth map. Then the \ib{pullback} of $E$ by $f$ is
the data of a smooth manifold
\[
f^*E = \set{(x, p)~:~ x \in X, p \in \pi\inv(f(x)) }
\]
along with the projection $p : f^*E \to M$ given by $(x,p) \mapsto x$, giving
$f^*E \to X$ the structure of a fiber bundle over $X$ with model fiber $F$.
The bundle $f^*E$ also comes equipped with a natural map $\alpha : f^*E \to E$
where $\alpha(x,p) = p$. Pullbacks give rise to the commutative diagram
\[\begin{tikzcd}
f^*E \ar[d, "p"']\ar[r, "\alpha"] & E \ar[d, "\pi"] \\
X \ar[r, "f"']& M
\end{tikzcd}\]
and are an instance of a general construction called the \ib{fiber product}.
\end{defn}
%
\begin{defn}
Let $\pi : E \to M$ be a fiber bundle. A \ib{local section} of $\pi : E \to M$
is a smooth map $\sigma : U \to E$ of an open set $U \subset M$ such that
$\pi \circ \sigma = \id_U$. If $U = M$, the section is called a \ib{global section}.
Equivalently, it is the smooth assignment of an element in $E_p$ to each point
$p \in U$. We denote the set of sections of $\pi : E \to M$ over a set $U$
as $\Gamma_U(E)$.
\end{defn}
%
A fiber bundle is a very general construction in which the fibers $F$ do
not necessarily admit extra structure. An important special case of a fiber bundle
is a vector bundle, where the fibers are vector spaces.
%
\begin{defn}
Let $M$ be a smooth manifold. A \ib{vector bundle} over $M$ is fiber bundle
$\pi : E \to M$ with model fiber $\R^n$ (or $\C^n)$ such that the local
trivializations $\varphi : \pi\inv(U) \to U \times \R^n$ (or $\C^n$)
restrict to linear isomorphisms on the fibers, i.e. for all $p \in U$, the restriction
$\varphi\vert_{\pi\inv(p)} : \pi\inv(p) \to \set{p} \times \R^n$
(or $\C^n$) is an isomorphism. The dimension $n$ of the fibers is called the
\ib{rank} of the vector bundle. A \ib{vector bundle homomorphism}
$\varphi : E \to L$ is a bundle homomorphism with the added stipulation that
the restrictions to the fibers $\varphi\vert_{E_x}$ are linear maps.
\end{defn}
%
\begin{exmp} \enumbreak
\begin{enumerate}
\item Given a smooth manifold $M$, the tangent bundle $TM = \coprod_{p \in M} T_pM$
is a vector bundle, where the projection $\pi : TM \to M$ is the mapping
$(p,v) \mapsto p$. The rank of $TM$ is the dimension of $M$.
Local coordinates on $(x^i)$ on $U \subset M$ induce maps
$\pi\inv(U) \to U \times \R^n$, where
$(p,v) \mapsto (x^1(p), \ldots x^n(p), v^1, \ldots v^n)$, where the $v^i$ are
the components of $v$ with respect to the coordinate bases $\partial_i$.
These maps define a basis for a topology on $TM$, which we then give
a smooth structure by declaring the projection $\pi : TM \to M$ to be
smooth. In addition, these maps define local trivializations for $TM$,
showing that it is a vector bundle.
\item A point in $\RP^n$ is a line $\ell$ in $\R^{n+1}$. The \ib{tautological
bundle} over $\RP^n$ is the bundle where the fiber over $\ell$ is the subspace
$\ell$ itself. An analogous construction can be done for Grassmannians
$\mathrm{Gr}_k(V)$, which parameterize $k$-dimensional subspaces of
a vector space $V$.
\qedhere\end{enumerate}
\end{exmp}
%
\begin{defn}
A \ib{Lie group} is a smooth manifold $G$ with a group structure such that the
multiplication map $(g,h) \mapsto gh$ and the inversion map $g \mapsto g\inv$
are smooth.
\end{defn}
%
Almost all of the groups we have discussed previously are Lie groups.
\begin{exmp} \enumbreak
\begin{enumerate}
\item The orthogonal groups $O_n$ and special orthogonal groups $SO_n$ are Lie groups.
\item The unitary groups $U_n$ and special unitary groups $SU_n$ are Lie groups.
\qedhere\end{enumerate}
\end{exmp}
%
Another important class of fiber bundles are principal bundles. Given a
Lie group $G$, the fibers of a principal $G$-bundle are $G$-torsors --
manifolds with a smooth free and transitive right (or left) $G$-action.
Given a $G$ torsor $M$ and a point $p$ in $M$, there is a diffeomorphism
$G \to M$, where $g \mapsto p \cdot g$. A $G$-torsor can be thought
of as $G$ without a distinguished identity element.
%
\begin{defn}
Let $G$ be a Lie group, and $M$ a smooth manifold. A \ib{principal } $G$\ib{-bundle}
over $M$ is the data of
%
\begin{enumerate}
\item A smooth manifold $P$ with a map $\pi : P \to M$.
\item A smooth right $G$-action on $P$ that preserves the fibers of $\pi$ and
is free and transitive on the fibers of $\pi$.
\item For every point $p \in M$, a neighborhood $U \subset M$ containing $p$ and
a $G$-equivariant diffeomorphism $\varphi: \pi\inv(U) \to U \times G$ (where
the right action on $U \times G$ is right multiplication on the second factor)
such that we get the commutative diagram
\[\begin{tikzcd}
\pi\inv(U) \ar[dr, "\pi"']\ar[rr, "\varphi"]&& U \times G \ar[dl, "p_U"]\\
& U
\end{tikzcd}\]
where $p_U$ denotes projection onto the first factor.
\end{enumerate}
A \ib{principal bundle homomorphism} is a bundle homomorphism $\varphi : P \to Q$
that is $G$-equivariant.
\end{defn}
%
\begin{exmp}
Given a smooth manifold $M$ and a point $p \in M$, a basis of the
tangent space is a linear isomorphism $b : \R^n \to T_pM$. The group
$GL_n\R$ acts freely and transitively on the set of bases $\mathcal{B}_p$ on the right by
$b \cdot g = b \circ g$. Then the \ib{frame bundle} of $M$, denoted
$\mathcal{B}(M)$ is the disjoint union
\[
\mathcal{B}(M) = \coprod_{p \in M}\mathcal{B}_p
\]
where $\pi$ is the projection map $(p,b) \mapsto p$. Like with the tangent
bundle, we use local coordinates on $M$ to give maps $\pi\inv(U) \to U \times GL_n\R$.
These maps can then be used to define a topology and smooth structure, and
form local trivializations of $\B(M)$, giving it the structure of a principal
$GL_n\R$-bundle over $M$.
\end{exmp}
%
\begin{exmp}
Given a smooth manifold $M$, a Riemannian metric $g$ induces an inner product
$g_p$ on each tangent space $T_pM$, where $g_p$ denotes the metric $g$ evaluated
at $p$. Then the set of orthonormal bases of $T_pM$ is the
set of all linear isometries $(\R^n, \langle\cdot,\cdot\rangle) \to (M, g_p)$
where $\langle\cdot,\cdot\rangle$ is the standard inner product on $\R^n$.
Then taking the disjoint union over all points $p\in M$ of orthonormal bases
for the tangent spaces $T_pM$ forms the \ib{orthonormal frame bundle}
$\mathcal{B}_O(M)$, which is a principal $O_n$ bundle.
\end{exmp}
%
\begin{defn}
Let $\pi : P \to M$ be a principal $G$-bundle over $M$, and let $F$ be a
smooth manifold with a smooth left $G$ action. Then the \ib{associated fiber
bundle}, denoted $P \times_G F$, is the set
\[
P \times_G F = P \times G / (p,g) \sim (p\cdot h, h\inv g)
\]
Since the group action on $P$ preserves the fibers, the projection
$p_1 : P \times F \to P$ composed with the projection $\pi : P \to M$
descends to the quotient, giving us a projection map
$\Phi : P \times_G F \to M$.
\end{defn}
%
The first thing to check is that $P \times_G F$ is a fiber bundle, justifying
the name.
%
\begin{prop}
Let $\pi : P \to M$ be a principal $G$-bundle and $F$ a manifold with a left $G$ action.
then the associated bundle $ \Phi : P \times_G F \to M$ is a fiber bundle with
model fiber $F$.
\end{prop}
%
\begin{proof}
We wish to provide local trivializations  $\Phi\inv(U) \to U \times F$ for the
associated bundle. Fix a local trivialization  $\psi : \pi\inv(U) \to U \times G$.
Then $\psi$ is of the form $\psi(p) = (\pi(p), \tilde{\psi}(p))$ for some
$\tilde{\psi} : \pi\inv(U) \to G$, satisfying $\tilde{\psi}(p\cdot g) = \tilde{\psi}(p)g$.
Define
\begin{align*}
\varphi : \Phi\inv(U) & \to U \times F                           \\
[p,f]                 & \mapsto (\pi(p), \tilde{\psi}(p)\cdot f)
\end{align*}
We first note that this is well defined on equivalence classes, since
\[
[p\cdot g, g\inv \cdot f] \mapsto (\pi(p \cdot g), \tilde{\psi}(p \cdot g)
\cdot g\inv \cdot f)
= (\pi(p), \psi(p)\cdot f)
\]
So $\varphi$ is well defined. We also note that $\Phi(\varphi[p,f]) = \pi(p)$
by how $\Phi$ was defined, so $\varphi$ is a local trivialization provided
it is a homeomorphism. To show this, we construct an inverse.
Define $\alpha : U \times F \to \Phi\inv(U)$
by $\alpha(u,f) = [\psi\inv(u,e), f]$, where $e$ denotes the identity element
of $G$. We then claim that $\alpha$ is the inverse. We compute
%
\begin{align*}
(\varphi \circ \alpha)(u,f) & = \varphi[\psi\inv(u,e),f] \\
                            & = (u, e\cdot f)            \\
                            & = (u,f)
\end{align*}
%
In the other direction, we compute
%
\begin{align*}
(\alpha \circ \varphi)[p,f] & = \alpha(\pi(p), \tilde{\psi}(p)) \\
                            & = [\psi\inv(\pi(p), e), f]        \\
                            & = [p, f]
\end{align*}
%
Therefore, $\varphi$ is a local trivialization, giving us that
$\Phi: P \times_G F \to M$ is a fiber bundle with model fiber $F$.
%
\end{proof}
%
There is a correspondence between sections of an associated bundle and
$G$-equivariant maps $P \to F$.
%
\begin{prop}
Let $\pi : P \to M$ be a principal $G$-bundle. Then there is a bijection
\[
\set{G\text{-equivariant maps } P \to F} \longleftrightarrow \Gamma_M(P \times_G F)
\]
\end{prop}
%
\begin{proof}
Since $F$ has a left $G$-action, we use it to define a right $G$-action
by $f \cdot g = g\inv \cdot f$. Then what we mean by a $G$-equivariant map is
a map $\varphi : P \to F$ such that
\[
\varphi(p\cdot g) = g\inv\cdot \varphi(p)
\]
We then wish to use $G$-equivariant map $\varphi$ to produce a section
$\tilde{\varphi} : M \to P \times_G F$ of the associated bundle. For a
point $x \in M$, pick any $p \in \pi\inv(x)$ in the fiber. Then define
\[
\tilde{\varphi}(x) = [p, \varphi(p)]
\]
where $[p, \varphi(p)]$ denotes the equivalence class of $(p, \varphi(p))$ in
$P \times_G F$. We first claim that this map is well defined, i.e. it is
independent of our choice of point $p \in \pi\inv(x)$. We know that $G$ acts
freely and transitively on $\pi\inv(x)$, so all points in the fiber are of
the form $p \cdot g$ for a unique $g \in G$. Then we have that for any $g \in G$
\[
(p\cdot g, \varphi(p \cdot g)) = (p \cdot g, g\inv \cdot \varphi(p)) \sim (p, \varphi(p))
\]
where we use the $G$-equivariance of $\varphi$ and the definition of the equivalence
relation on the associated bundle. In addition, this induced map is
a section of $P \times_G F \to M$, since the image is represented by an
element of $P \times F$ with an element of the fiber $\pi\inv(x)$ in the first
factor. \\

Conversely, given a section $\sigma : M \to P \times_G F$, we wish to produce
a $G$-equivariant map $P \to F$. Given such a section, and a point $x \in M$,
we have that $\sigma(x) = [p, f]$ for some $p \in P$ and $f \in F$. Then
define $\tilde{\sigma} : P \to F$ such that $\tilde{\sigma}(p) = f$, and
$\tilde{\sigma}(p \cdot g) = g\inv \cdot f$. Since $G$ acts freely and transitively
on the fibers, this specifies the map on every point of $P$. In addition,
$\tilde{\sigma}$ is $G$-equivariant by construction.
Then the maps
\[
\set{G\text{-equivariant maps } P \to F} \longleftrightarrow \Gamma_M(P \times_G F)
\]
we provided are easily verified to be inverses to each other, giving us the
correspondence.
\end{proof}
%
In some sense, the geometry of the associated fiber bundle $P \times_G F$
is controlled by the group $G$, as $G$ determines a distinguished
group of symmetries of the fiber $F$. Given an element $p \in P$, the element
$p$ determines a diffeomorphism $\varphi_p : F \to (P \times_G F)_{\pi(p)}$ via the
mapping $f \mapsto [p,f]$, and the group action of $G$ on $P$ gives a
transformation law on the associated bundle via the group action of $G$ on $F$.
%
\begin{exmp}[The tangent bundle as an associated bundle]
Given a manifold $M$, we can take the frame bundle $\pi : \B(M) \to M$, which is a
principal $GL_n\R$ bundle. The group $GL_n\R$ acts linearly on $\R^n$ in the
standard way, giving us an associated vector bundle $\B(M) \times_{GL_n\R} \R^n$.
We claim that this bundle is isomorphic to the tangent bundle $TM$, i.e. there
exists an diffeomorphism $\varphi : \B(M) \times_{GL_n\R} \R^n \to TM$ that
restricts to linear isomorphisms on the fibers and the diagram
\[\begin{tikzcd}
\B(M) \times_{GL_n\R} \R^n \ar[rr, "\varphi"] \ar[dr] && TM \ar[dl] \\
& M
\end{tikzcd}\]
commutes, where the maps to $M$ are the bundle projections. Recall that
elements of $\B(M) \times_{GL_n\R} \R^n$ are represented by pairs $(b, v)$,
where $b: \R^n \to T_{\pi(b)}M$ is a linear isomorphism, and $v$ is a vector in $\R^n$.
Then define $\varphi$ by
\[
 \varphi[b,v] = (\pi(b), b(v))
\]
This is well defined, since $\varphi[b\circ g, g\inv(v)] = (\pi(b \circ g),
(b\circ g)(g\inv(v))) = (\pi(b), b(v))$. This is an isomorphism, where the
inverse mapping maps $(p,v) \in TM$ to $(b, \tilde{v})$ where $b$ is any
basis of $T_pM$ and $\tilde{v}$ is the coordinate representation of $v$ in the
basis $b$. In this example, we see that the associated bundle identifies the same vector
under different coordinate transformations, which defines the symmetries of
$TM$.
\end{exmp}
%
In a similar fashion, the cotangent bundle $T^*M$ to a manifold $M$ is
naturally isomorphic to the associated bundle $\B(M) \times_{GL_n\R} (\R^n)^*$,
where $GL_n\R$ acts on $(\R^n)^*$ by the dual of the standard representation
(i.e. $A \in GL_n\R$ acts by $(A\inv)^*$.) \\
%

In general, given a rank $n$ vector bundle $E \to M$, we can construct the
frame bundle  $\B(E)$ for $E$, where the fiber over $x \in M$ is the
$GL_n\R$-torsor of bases for the fiber $E_x$. We can then
recover $E$ by taking the associated bundle $\B(E) \times_{GL_n\R} \R^n$,
so the process of taking frames and constructing associated bundles are inverses
of each other. In addition, if $E$ is an associated vector bundle
$P \times_G V$ obtained via a linear representation $\rho : G \to GL(V)$, the
elements of $P$ have a natural interpretation as bases of the fibers of $E$.
In this case, each element $p \in P$ determines a linear isomorphism
$\varphi_p : V \to E_p$, giving a $G$-equivariant $\varphi : P \to \B(E)$ of
$P$ to the frame bundle of $E$, where $G$ acts on $\B(E)$ through the
representation $\rho$.
%
\begin{defn}
 Let $\pi : P \to M$ be a principal $G$-bundle, and $\rho : H \to G$ a group
 homomorphism. The map $\rho$ gives $G$ a left $H$ action where
 $h \cdot p = \rho(h) \cdot p$. A \ib{reduction of structure group} is the
 data of a principal $H$ bundle $\varphi : Q \to M$ and an $H$-equivariant bundle
 homomorphism $F : Q \to P$.
\end{defn}
%
The map $F : Q \to P$ induces a map $\tilde{F} : Q \times_H G \to P$, where we
map the equivalence class $[q,g]$ to $F(q) g$. This is well defined
on equivalences classes since
%TODO this next line might be unnecessary, since it is the same map
%constructed using the correspondence with sections -- double check this.
\[
 (q \cdot h, \rho(h)\inv g) \mapsto F(q \cdot h)\rho(h)\inv g = F(q)\rho(h)\rho(h)\inv g
 = F(q)g
\]
%
\begin{exmp}[Reduction from $GL_n\R$ to $O_n$]
 Let $M$ be a smooth manifold, and $\pi : \B(M) \to M$ its bundle of frames.
 The inclusion map $\iota: O_n \hookrightarrow GL_n\R$ gives an action of $O_n$ on
 $\B(M)$, where given $b \in \B(M)$ and $T \in O_n$, $b \cdot T = b \cdot \iota(T)$.
 We then take the quotient by this $O_n$ action, giving us a quotient map
 $q : \B(M) \to \B(M) / O_n$. Since the inclusion is injective, the $O_n$ action
 is free on $\B(M)$, so this gives $q : \B(M) \to \B(M) / O_n$ the structure of a
 $O_n$ bundle. In addition, the action of $O_n$ preserves the fibers of
 $\pi : \B(M) \to M$, so $\pi$ descends to the quotient, so $\B(M) / O_n \to M$
 is a fiber bundle over $M$ with model fiber $GL_n\R / O_n$. Since $GL_n\R$ deformation
 retracts onto $O_n$ via the Gram-Schmidt algorithm, $GL_n\R / O_n$ is
 contractible, so the fiber bundle $\B(M)/ O_n \to M$ admits global sections.
 Then given a section $\sigma : M \to \B(M) / O_n$, this gives an $O_n$ bundle
 over $M$ via the pullback $\sigma^*\B(M)$. In addition, we get a $O_n$-equivariant
 map $\sigma^*\B(M) \to \B(M)$ given by $(p, b) \mapsto (p, \iota(b))$. The
 bundle $\sigma^*\B(M)$ can be thought of as the bundle of orthonormal frames
 with respect to a Riemannian metric on $M$, and the fact that $\B(M) / O_n$
 admits sections corresponds to the fact that every manifold admits a Riemannian
 metric.
\end{exmp}
%
Principal bundles can be thought of as a generalization of covering spaces.
For discrete groups $G$, the data of a principal $G$-bundle $P \to M$ is the
data of a (possibly disconnected) cover of $M$ with a deck transformation
group isomorphic to $G$ acting freely and transitively on the fibers. If
$G$ is finite, this is a $|G|$-fold covering, and if $G$ is countable, the
cover is an infinitely sheeted cover. Discreteness of $G$ allows for a very
complete description of all principal $G$-bundles over a given manifold $M$.
%
\begin{defn}
 Let $(M,x_0)$ be a pointed topological space (in our case, usually a manifold),
 i.e. a space $M$ with a choice of distinguished basepoint $x_0 \in M$. Then
 a \ib{pointed principal $G$-bundle} over $M$ is the data of a principal $G$-bundle
 $\pi : P \to M$, along with a choice of basepoint $p_0 \in \pi\inv(x_0)$. We
 denote this as $\pi : (P,p_0) \to (M, x_0)$. We denote the set of isomorphism
 classes of pointed principal $G$-bundles over $(M,x_0)$ as $\Bun_G(M ,x_0)$.
\end{defn}
%
We now recall some properties of covering spaces. A pointed covering space
$\pi : (\tilde{M}, \tilde{x_0}) \to (M, x_0)$ has the \ib{path lifting} property,
i.e. given a path $\gamma : I \to M$ with $\gamma(0) = x_0$, there exists a
unique path $\tilde{\gamma} : I \to \tilde{M}$ with $\gamma(0) = \tilde{x_0}$
such that $\pi \circ \tilde{\gamma} = \gamma$. In addition, the covering
space has the \ib{homotopy lifting property}. Given paths
$\gamma_1, \gamma_2 : I \to M$ with $\gamma_i(0) = x_0$ and a homotopy
$F : I \times I \to M$ from $\gamma_1$ to $\gamma_2$, there exists a unique
homotopy $\tilde{F} : I \times I \to \tilde{M}$ between $\tilde{\gamma}_1$
and $\tilde{\gamma_2}$. These two facts define an action of $\pi_1(M, x_0)$
on the fiber $\pi\inv(x_0)$ where given a homotopy class $[\gamma]$ of
loops, the action of $[\gamma]$ on $p \in \pi\inv(x_0)$ is the endpoint
of $\tilde{\gamma}(1)$ of the lifted loop $\tilde{\gamma}$. The unique
path lifting and homotopy lifting properties guarantees that this defines
a group action, called the \ib{holonomy} of the covering space. In the
special case that our covering space is a principal $G$-bundle with a discrete $G$,
the point $\tilde{\gamma}(1)$ is equal to $\tilde{x_0} \cdot g$ for a unique
group element $g \in G$, so the holonomy action determines a group homomorphism
$\pi_1(M, x_0) \to G$. In fact, the holonomy completely determines the bundle.
%
\begin{thm}
 For a discrete group $G$ and a connected manifold $M$, there is a bijection
 \[
  \Bun_G(M,x_0) \longleftrightarrow \hom(\pi_1(M,x_0), G)
 \]
\end{thm}
%
\begin{proof}
 We once more provide maps in both directions. Every pointed principal $G$-bundle
 $(P,p_0) \to (M, x_0)$ determines a group homomorphism $\pi_1(M,x_0) \to G$
 by the holonomy action. In the other direction, given a group homomorphism
 $\varphi : \pi_1(M, x_0) \to G$, we want to construct a pointed principal $G$
 bundle $(P, p_0) \to (M, x_0)$ with holonomy $\varphi$. Every connected pointed manifold
 $(M, x_0)$ has a \ib{universal cover} $(\tilde{M}, \tilde{x}_0) \to (M,x_0)$,
 which has a free and transitive action by $\pi_1(M,x_0)$, making it a pointed
 principal $\pi_1(M,x_0)$ bundle over $(M,x_0)$. In addition, the homomorphism
 $\varphi : \pi_1(M, x_0) \to G$ determines a left action of $\pi_1(M,x_0)$
 on $G$ by $[\gamma] \cdot g = \varphi[\gamma] g$. This allows us to define
 the associated bundle
 \[
  P = \tilde{M} \times_{\pi_1(M, x_0)} G = \tilde{M} \times G
  / (x,g) \sim (x \cdot [\gamma], \varphi[\gamma]\inv g)
 \]
 where our distinguished basepoint $p_0 \in P$ is $p_0 = [\tilde{x}_0, e]$.
 This is a fiber bundle with model fiber $G$, and the right action of $G$
 on $\tilde{M} \times G$ descends to a free an transitive action on the quotient,
 making it a principal $G$-bundle. We then check the holonomy action of a
 loop $[\gamma] \in \pi_1(M, x_0)$. We compute for $p$ in the fiber over $x_0$
 \[
  p \cdot [\gamma] = [\tilde{x_0} \cdot [\gamma], h] \sim [x_0, \varphi[\gamma]\cdot h]
 \]
 So the holonomy is $\varphi$, giving us the desired correspondence.
\end{proof}
%
The set $\Bun_G(M, x_0)$ comes equipped with an action of $G$, where the action
of $g \in G$ on a pointed bundle $(P, p_0) \to (M, x_0)$  is by permuting basepoints,
i.e. $(P, p_0) \cdot g = (P, p_0 \cdot g)$, where the covering map is left untouched.
Quotienting by this group action gives us the identification of $\Bun_G(M,x_0) / G$
with isomorphism classes of principal $G$-bundles $P \to M$. where the
isomorphisms are without basepoints. From the perspective of pointed bundles
$(P,p_0) \to (M, x_0)$ as homomorphisms $\varphi : \pi_1(M, x_0) \to G$,
the group action of $G$ on $\Bun_G(M,x_0)$ is an action by conjugation.
Given $(P, p_0) \to (M, x_0)$ with holonomy $\varphi : \pi_1(M, x_0) \to G$,
we want to know the holonomy $\varphi_g$ of the bundle
$(P, p_0 \cdot g) \to (M, x_0)$. Given a loop $\gamma$ based at $x_0$, let
$\tilde{\gamma}$ denote its lift to $(P, p_0)$ by uniqueness of path lifting,
the lift of $\gamma$ to $(P, p_0 \cdot g)$ is the path
$\tilde{\gamma}_g(t) = \tilde{\gamma}(t) \cdot g$. Then
%
\begin{align*}
 \tilde{\gamma_g(1)} & = \tilde{\gamma}(1) \cdot g                             \\
                     & = p_0 \cdot \varphi[\gamma] \cdot g                     \\
                     & = p_0 \cdot g \cdot g\inv \cdot \varphi[\gamma] \cdot g
\end{align*}
%
So $\varphi_g = g \inv\cdot\varphi\cdot g$.
%
